---
title: "transmit uplink diags"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---
# notes

# setup instructions
edit `data_search_path` to where you're wildlife computers portal downloaded data is and run this. 

the r code below expects `\*-Status.csv`, `\*-All.csv`, `*-Corrupt.csv` files and does its best to recover from errors and missing data. it should be robust to several different file storage solutions, searches recursively in the data directory for all status and all files and should ignore all other files in the path.

if you have a metadata csv file with columns for `ptt` and `deploydate_yyyymmdd` you can set `metadata_path` to the csv and set `deploytime_metadata` to `TRUE`. this will cut off any undeleted data on the tag prior to deployment which sometimes makes the graphs messy and hard to read and impacts tx rate estimates etc.

```{r setup-change-this-path, echo = FALSE}
# change this path to your tag archives
data_search_path <- ""

# set deploytime_metadata to TRUE and set your path to your metadata csv
# if you want to cut off predeployment data
# metadata_path must point to a csv
# must have columns ptt and deploydate_yyyymmdd (and expects a date in that format)
deploytime_metadata <- FALSE
metadata_path <- ".csv"

```

```{r setup, echo = FALSE}
# helper functions
dateseq <- function(d, hours = FALSE) {
  unit <- 60 * 60 * 24
  if (hours) 
    unit <- 60 * 60
  mind <- min(d, na.rm = TRUE)
  maxd <- max(d, na.rm = TRUE)
  std <- ceiling(mind/unit) * unit
  end <- ceiling(maxd/unit) * unit
  seq(std, end, by = unit)
}

matchtimes <- function(t1, t2) {
  # t1, t2 are numeric
  findInterval(t1, c(-Inf, head(t2, -1)) + c(0, diff(t2)/2))
}
```

# plots
```{r load-and-calc, echo = FALSE}
# load metadata if required
if(deploytime_metadata) {
  # load metadata and make a deploytimes cutoff
  deploymeta <- read.table(metadata_path, header = TRUE, sep = ',', stringsAsFactors = FALSE)
  # make a data cutoff for each tag to remove extraneous data
  deploytimes <- as.POSIXct(
	  as.character(deploymeta$deploydate_yyyymmdd),
	  format = "%Y%m%d",
	  tz = "UTC"
)

# make datenum
deploytimes <- as.numeric(deploytimes)
}

# look for status and all data files
fpaths <- list.files(data_search_path, recursive = TRUE)
dese <- grepl("*-Status.csv$|*-All.csv$|*-Corrupt.csv$", fpaths)

# if there is nothing then throw an error
if(all(!dese)) stop(paste("I can't find any data in", data_search_path))

# pair down the files to just these
fpaths <- fpaths[dese]
fnames <- basename(fpaths)

# grab the ptt or whatever is the prefix so can match sta and all
ptts <- sub("-Status.csv|-All.csv|-Corrupt.csv", "", fnames)
uptts <- unique(ptts)
nptts <- length(uptts)

# set up a list to hold kable data
kables <- list()
kables_uplinks_per_day <- list()
kables_corrupt <- list()

# iterate through and check the transmits
for(i in 1:nptts) {
  curptt <- uptts[i]
  dese <- fpaths[ptts == curptt]
  
  # should never trip this
  if(length(dese) == 0) stop(paste("something has gone horribly wrong at index",i, "prefix", curptt))
  
  # find the paths for status and all
  sta_fpath <- dese[grepl("-Status.csv$", dese)]
  all_fpath <- dese[grepl("-All.csv$", dese)]
  cor_fpath <- dese[grepl("-Corrupt.csv$", dese)]
  
  # throw a warning if there isn't an all
  # this is weird there should always be an all
  # otherwise load the file and grab the uplinks
  curall <- NULL
  if(length(all_fpath) == 0) {
    warning(paste("there is no all for", curptt, "but also weird shoudn't happen ever...?"))
  } else {
    # load file
      tryCatch({
    curall <- read.table(file.path(data_search_path, all_fpath), header = TRUE, sep = ',', stringsAsFactors = FALSE)
      }, 
        error = function(e) {
          warning(e)  
        },
        warnings = function(e) {
          warning(e)
        },
      finally = {})
  }
  
    # do a little bit if we have a file
    if(!is.null(curall)) {
      # make a bunch of different dates
      curall$date_posix <- as.POSIXct(curall$Msg.Date, format = "%m/%d/%Y %H:%M:%S", tz = 'utc')
      curall$datenum <- as.numeric(curall$date_posix)
      curall$day <- format(curall$date_posix, "%Y%m%d")
      
      # sort by date just in case sometimes the order gets messed up
      curall <- curall[order(curall$datenum), ]
      
      # cutoff
      if(deploytime_metadata) {
        curall <- curall[curall$datenum >= deploytimes[deploymeta$ptt == curall$Platform.ID.No.[1]], ]
      }
      
      # uplinks per day
      uplinks_per_day <- as.data.frame(table(curall$day))
      names(uplinks_per_day) <- c(paste(curall$DeployID[1], curptt), 'count')
      kables_uplinks_per_day[[i]] <- uplinks_per_day
      
      # cumulative unlinks plot
      plot(curall$datenum, 1:nrow(curall), xaxt = 'n', main = paste(curall$DeployID[1], curall$Platform.ID.No[1], sep = ' / '), xlab = "", ylab = "rx", las = 1)
      lines(curall$datenum, 1:nrow(curall))
      
      # human readible axis
      ax <- dateseq(curall$datenum)
      axis(1, at = ax, lab = as.POSIXct(ax, tz = 'utc', origin = "1970-01-01"), las = 2)
    }

  cursta <- NULL
  if(length(sta_fpath) == 0) { 
    warning(paste("there is no status for", curptt))
  } else {
    # load file
        tryCatch({
      cursta <- read.table(file.path(data_search_path, sta_fpath), header = TRUE, sep = ',', stringsAsFactors = FALSE)
        }, 
        error = function(e) {
          warning(e)  
        },
        warnings = function(e) {
          warning(e)
        },
      finally = {})
  }
  
  # if we have a file do a little bit
  if(!is.null(cursta)) {
    # make a real dates
    cursta$date_posix <- as.POSIXct(cursta$Received, format = "%H:%M:%S %d-%b-%Y", tz = "UTC")
    cursta$datenum <- as.numeric(cursta$date_posix)
    
    # sort just in case
    cursta <- cursta[order(cursta$datenum), ]
    
    # cutoff
    if(deploytime_metadata) {
      cursta <- cursta[cursta$datenum >= deploytimes[deploymeta$ptt == cursta$Ptt[1]], ]
    }
    
    # filter out non crc messages as they tend to be mad corrupt
    # make sure there aren't NAs because that'll mess up the filter
    cursta$Type[is.na(cursta$Type)] <- ""
    cursta <- cursta[cursta$Type == 'CRC', ]
  
  if(nrow(cursta) > 0 ) {
   cursta2 <- cursta[!is.na(cursta$Transmits), ]  
   cursta2$difftime <- c(NA, diff(cursta2$datenum))/60/60/24 # in days
   cursta2$difftransmits <- c(NA, diff(cursta2$Transmits))
   cursta2$transmits_per_day <- cursta2$difftransmits / cursta2$difftime
   
   # cumulative transmits plot
   plot(cursta$datenum, cursta$Transmits, xaxt = 'n', main = paste(cursta$DeployID[1], cursta$Ptt[1], sep = ' / '), xlab = "", ylab = "tx", las = 1)
   lines(cursta$datenum[!is.na(cursta$Transmits)], cursta$Transmits[!is.na(cursta$Transmits)])
   
   # human readible axis
   ax <- dateseq(curall$datenum)
   axis(1, at = ax, lab = as.POSIXct(ax, tz = 'utc', origin = "1970-01-01"), las = 2)
   
   # plot on transmits per day
   # make a midpoint
   xx <- c(NA, rowSums(cbind(cursta2$datenum[2:nrow(cursta)], cursta2$datenum[1:(nrow(cursta) - 1)]))/2)
   yy <- c(NA, rowSums(cbind(cursta2$Transmits[2:nrow(cursta)], cursta2$Transmits[1:(nrow(cursta) - 1)]))/2)
   text(xx, yy, paste(round(cursta2$transmits_per_day), 'tx/day'))
   
  } else {
    cursta <- NULL
    warning(paste("no CRC'd status messages for", curptt))
  }
}
  
  if(!is.null(curall) & !is.null(cursta)) {
    # plot uplinks against transmits perday
    if(nrow(cursta2) > 1) {
    cursta2$nuplinks <- NA
    
    # merge in number of uplinks for same time period
    for(p in 2:nrow(cursta2)) {
      st <- cursta2$datenum[p - 1]
      en <- cursta2$datenum[p]
      
      dese <- curall$datenum >= st & curall$datenum <= en
      cursta2$nuplinks[p] <- length(which(dese))
    }
    
    cursta2$uplinks_per_day <- cursta2$nuplinks / cursta2$difftime
    cursta2$uplink_transmits_ratio <- cursta2$uplinks_per_day / cursta2$transmits_per_day
    
    # save a table
    kables[[i]] <- cursta2[, c('DeployID', 'Ptt', 'date_posix', 'datenum', 'difftime', 'Transmits', 'difftransmits', 'transmits_per_day', 'nuplinks', 'uplinks_per_day', 'uplink_transmits_ratio')]
    }
  }
  
  # load the corrupt file if it is there
  curcor <- NULL
  if(length(cor_fpath) == 0) { 
    warning(paste("there is no corrupt for", curptt))
  } else {
    # load file
        tryCatch({
      curcor <- read.table(file.path(data_search_path, cor_fpath), header = TRUE, sep = ',', stringsAsFactors = FALSE)
        }, 
        error = function(e) {
          warning(e)  
        },
        warnings = function(e) {
          warning(e)
        },
      finally = {})
  }
  
  # if we have a file do a little bit
  if(!is.null(curcor)) {
    # make a real dates
    curcor$date_posix <- as.POSIXct(curcor$Date, format = "%H:%M:%S %d-%b-%Y", tz = "UTC")
    curcor$datenum <- as.numeric(curcor$date_posix)
    
    # sort just in case
    curcor <- curcor[order(curcor$datenum), ]
    
    # cutoff
    if(deploytime_metadata) {
      curcor <- curcor[curcor$datenum >= deploytimes[deploymeta$ptt == curcor$Ptt[1]], ]
    }
    
    # count
    yy <- 1:nrow(curcor)
    curcor$yy <- yy
    
    plot(curcor$datenum, yy, xaxt = 'n', main = paste(curcor$DeployID[1], curcor$Ptt[1], sep = ' / '), xlab = "", ylab = "no. corrupt", las = 1)
    lines(curcor$datenum, yy)
    ax <- dateseq(curcor$datenum)
    axis(1, at = ax, lab = as.POSIXct(ax, tz = 'utc', origin = "1970-01-01"), las = 2)
  }
  
  kables_corrupt[[i]] <- curcor
}
```

# all corrupt
```{r plot-all-corrupt, echo = FALSE}
xx <- lapply(kables_corrupt, function(x) (x$datenum - min(x$datenum)) / 60 / 60 ) # hours
yy <- lapply(kables_corrupt, '[[', 'yy')

plot(do.call(c, xx), do.call(c, yy), type = 'n', xlab = "running hours", ylab = "no. corrupt")

for(i in 1:length(kables_corrupt)) {
  lines(xx[[i]], yy[[i]], col = i)
}

plot(0, 0, type ='n', xlab = '', ylab = '', axes = FALSE)
legend("center",  col = 1:length(kables_corrupt), lty = 1, legend = sapply(kables_corrupt, function(x) paste(x$DeployID[1], x$Ptt[1])), border = FALSE)
```

# histos: rx per day, tx per day, rx/tx ratio
purple broken line = mean
annotation = mean (sd)
```{r hists-uplinks-per-day, echo = FALSE, results = 'asis'}
for(i in 1:length(kables_uplinks_per_day)) {
  titlelab <- names(kables_uplinks_per_day[[i]])[1]
  
  par(mfrow = c(2, 2))
  
  if(!is.null(kables_uplinks_per_day[[i]])) {
    hist(kables_uplinks_per_day[[i]]$count, main = titlelab, xlab = 'rx/day')
    
    uppdbar <- mean(kables_uplinks_per_day[[i]]$count, na.rm = TRUE)
    uppdsd <- sd(kables_uplinks_per_day[[i]]$count, na.rm = TRUE)
    
    abline(v = uppdbar, col = "purple", lty = 2)
    text(uppdbar, .75, paste0(round(uppdbar, 1), " (", round(uppdsd, 2), ")"), col = "purple")
  } else {
    plot(0, 0, type = 'n', axes = FALSE, xlab = '', ylab = '')
    text(0, 0, "rx/day: na")
  }
  
  if(!is.null(kables[[i]])) {
    txpd <- kables[[i]]$transmits_per_day[which(kables[[i]]$difftime > 0.5 & kables[[i]]$difftime < 5)]
    
    if(length(txpd) > 0) {
      txpdbar <- mean(txpd, na.rm = TRUE)
      txpdsd <- sd(txpd, na.rm = TRUE)
    
      hist(txpd, main = '', xlab = 'tx/day')
      abline(v = txpdbar, col = "purple", lty = 2)
      text(txpdbar, .75, paste0(round(txpdbar), " (", round(txpdsd), ")"), col = "purple")
    } else {
      plot(0, 0, type = 'n', axes = FALSE, xlab = '', ylab = '')
      text(0, 0, "tx/day: na")
    }
  } else {
    plot(0, 0, type = 'n', axes = FALSE, xlab = '', ylab = '')
    text(0, 0, "tx/day: na")
  }
  
  if(!is.null(kables[[i]])) {
    upptx <- kables[[i]]$uplink_transmits_ratio[which(kables[[i]]$difftime > 0.5 & kables[[i]]$difftime < 5)]
    
    if(length(upptx) > 0) {
      upptxbar <- mean(upptx, na.rm = TRUE)
      upptxsd <- sd(upptx, na.rm = TRUE)
    
      hist(upptx, main = '', xlab = 'rx/tx')
      abline(v = upptxbar, col = "purple", lty = 2)
      text(upptxbar, .75, paste0(round(upptxbar, 2), " (", round(upptxsd, 2), ")"), col = "purple")
    } else {
      plot(0, 0, type = 'n', axes = FALSE, xlab = '', ylab = '')
      text(0, 0, "rx/tx: na")
    }
  } else {
    plot(0, 0, type = 'n', axes = FALSE, xlab = '', ylab = '')
    text(0, 0, "rx/tx: na")
  }
}
```

# tables: rx/day
```{r print-uplinks-per-day, echo = FALSE, results = 'asis'}
for(i in 1:length(kables_uplinks_per_day)) {
  print(knitr::kable(kables_uplinks_per_day[[i]]))
  cat("\n")
}
```

# tables: rx/tx
```{r print-ratio, echo = FALSE, results = 'asis'}
for(i in 1:length(kables)) {
  print(knitr::kable(kables[[i]]))
  cat("\n")
}
```

